Community biomass and species richness temporal trends reveal contrasting assembly and disassembly processes in natural food-webs

### 2024-08-06 paper/appendix.Rmd
Finalise manuscript preparation
Add redundancy + update fig & tables

```{r, echo=FALSE}
knitr::opts_chunk$set(
  cache = FALSE,
  collapse = TRUE,
  comment = "#>",
  #fig.dim = c(7, 7),
  fig.align = "center",
  fig.fullwidth = TRUE,
  fig.show = "hold",
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = TRUE
)
options(
  #kableExtra.latex.load_packages = FALSE,
  knitr.table.format = "latex"
)
```

@@

theme_set(mytheme)
```
```{r}
var_analysis <- c("ct_ff", "w_trph_lvl_avg", "log_rich_std", "log_bm_std")
```

\beginsupplement
\newpage

```{r}
tar_load(tab_inla_r2_rand)
kable(tab_inla_r2_rand,
  row.names = FALSE,
  booktabs = T,
  caption = "Marginal and Conditional R-squared for the INLA models estimating
  the temporal trends of the food-web structure. The table also displays the
  standard deviation associated to the random effects and the general error of
  the models. Confidence interval at 95\\% are displayed between square brackets.",
  label = "inlarsq") %>%
  collapse_rows(columns = c(1, 2, 3), valign = "top") %>%
  kable_styling(latex_options = c("scale_down"))
```

@@

@@

```{r}
tar_load(r2_sem_tab)
kable(r2_sem_tab,
  row.names = FALSE,
  booktabs = T,
  caption = "Marginal and conditional R squared for the Structural Equation
  Models.",
  label = "r2sem",
  ) %>%
  kableExtra::collapse_rows(columns = 1, valign = "top")
```

```{r}
tar_load(c(pca_table, op_table, com_metric_tab))
tab_quanti <- purrr::map_dfr(
  list(com_metric_tab, op_table, pca_table),
  function(x) {colnames(x)[1] <- "Variable"; x}
)
#purrr::map_int(list(com_metric_tab, op_table, pca_table), nrow)

tab_quanti %>%
  kbl(
    caption = "Summary distribution of community metrics, fish monitoring and of
    environmental variables used in the PCA for the spatial SEM.",
    booktabs = T
    ) %>%
  kable_styling() %>%
  pack_rows("Community metrics", 1, 8) %>%
  pack_rows("Summary of fish monitoring", 9, 12) %>%
  pack_rows("Summary of Envionmental variables", 13, 26)
```

>>>>

```{r, echo=FALSE}
knitr::opts_chunk$set(
  cache = FALSE,
  collapse = TRUE,
  comment = "#>",
  #fig.dim = c(7, 7),
  fig.align = "center",
  fig.fullwidth = TRUE,
  fig.pos = "H",
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = TRUE
)
options(
  #kableExtra.latex.load_packages = FALSE,
  knitr.table.format = "latex",
  knitr.graphics.rel_path = FALSE
)
```

@@

theme_set(mytheme)
```

\beginsupplement

# Supplementary figures

@@

```{r}
sem_sp_mod_cap <- "Direct effects of biomass and species richness on food-web
structure, derived from the structural equation models for the (a) spatial and
(b) bioenergetic model. Points display the data, black lines display the
predicted relationships."

```

```{r, fig.cap=sem_sp_mod_cap, out.width = "100%"}
tar_load(p_sem_sp_sim_file)
knitr::include_graphics(p_sem_sp_sim_file)
```

\pagebreak

# Supplementary tables

```{r}
tar_load(r2_inla_tab)
kable(r2_inla_tab,
  row.names = FALSE,
  booktabs = T,
  caption = "Marginal and Conditional R-squared for the INLA models estimating
  the temporal trends of the food-web structure. Mean [95\\% Credible Interval].",
  label = "inlarsq") %>%
  kable_styling()
```

```{r}
tar_load(c(op_table, com_metric_tab))
tab_quanti <- purrr::map_dfr(
  list(com_metric_tab, op_table),
  function(x) {colnames(x)[1] <- "Variable"; x}
)
tab_quanti %>%
  kbl(
    caption = "Summary distribution of community metrics and fish monitoring variables.",
    booktabs = T
    ) %>%
  kable_styling() %>%
  pack_rows("Community metrics", 1, 8) %>%
  pack_rows("Summary of fish monitoring", 9, 13)
```

```{r}
tar_load(tab_inla_rand)
tab_inla_rand %>%
  mutate(Response = factor(Response, levels = unique(var_replacement()))) %>%
  arrange(Response) %>%
  kable(
    row.names = FALSE,
    booktabs = T,
    caption = "Standard deviation associated to the random effects and the residual error of
    the models. Mean [95\\% Credible Interval].",
    label = "inlarand") %>%
  collapse_rows(columns = c(1), valign = "top") %>%
  kable_styling()
```

```{r}
tar_load(semeff_tot_tab)
ti <- semeff_tot_tab %>%
  pivot_wider(names_from = "Effect type", values_from = "Effect") %>%
  mutate(Response = factor(Response, levels = unique(var_replacement()))) %>%
  arrange(Model, Response)
kable(ti %>%
    select(-Model),
  row.names = FALSE,
  booktabs = T,
  caption = "Direct and total Standardized effects derived from the structural
  equation models. [95\\% Confidence Interval]. The confidence intervals were
  obtained with bootstrap (see Methods).") %>%
  collapse_rows(columns = c(1), valign = "top") %>%
  pack_rows(index = table(ti$Model))
```

```{r}
tar_load(r2_sem_tab)
kable(r2_sem_tab %>%
  mutate(
    Type = factor(Type, levels = effect_type_var()),
    Response = factor(Response, levels = unique(var_replacement()))
    ) %>%
  arrange(Type, Response),
  row.names = FALSE,
  booktabs = T,
  caption = "Marginal and conditional R squared for the Structural Equation
  Models.",
  label = "r2sem",
  ) %>%
  kableExtra::collapse_rows(columns = 1, valign = "top")
```

@@

### 2024-08-06 paper/manuscript.Rmd
Add redundancy + update fig & tables

```{r, results=FALSE}
tar_load(glob_tps_trends_decade_no_drivers)
gbm <- get_effect_ci(resp = "log_bm_std", ci_lvl = "level:0.80", p = TRUE)
gr <- get_effect_ci(resp = "log_rich_std", ci_lvl = "level:0.80", p = TRUE)
gt <- get_effect_ci(resp = "w_trph_lvl_avg", ci_lvl = "level:0.80", p = FALSE)
gct <- get_effect_ci(resp = "ct_ff", ci_lvl = "level:0.80", p = FALSE)
```

@@

```{r}
dfv <- function(
  x = summary_pred_station_tps10,
  r = "ct_ff",
  v = "min",
  d = 4
) {
  out <- x[x$response == r, ][[v]]
  format(round(out, d), nsmall = d)
}
grbmn <- dfv(r = "log_bm_std", v = "min", d = 0)
grbmx <- dfv(r = "log_bm_std", v = "max", d = 0)
grrhn <- dfv(r = "log_rich_std", v = "min", d = 0)
grrhx <- dfv(r = "log_rich_std", v = "max", d = 0)
grctn <- dfv(r = "ct_ff", v = "min", d = 2)
grctx <- dfv(r = "ct_ff", v = "max", d = 2)
grtln <- dfv(r = "w_trph_lvl_avg", v = "min", d = 2)
grtlx <- dfv(r = "w_trph_lvl_avg", v = "max", d = 2)
```

Using a Bayesian hierarchical model to jointly estimate the temporal trends of
food-web structure at all sites, we found weak evidence of an overall decrease
in community biomass (CI80%:`r gbm` by decade; median [Credible Interval], Fig.
S1) but a weak evidence of an overall temporal increase in species richness
(CI80%: `r gr` by decade) across the `r nb_st` sites. We did not find evidence
of average temporal trends in connectance and average trophic level (CI80%:
`r gct` and `r gt` respectively). However, we found a great heterogeneity in the
temporal trends estimated at the site level (community biomass: from `r grbmn`%
to `r grbmx`% per decade, species richness: `r grrhn` to +`r grrhx`%, average
trophic level: `r grtln` to +`r grtlx`, connectance: `r grctn` to +`r grctx`,
Fig. S1) which was well captured by the hierarchical model with conditional R2
going from `r r2cond$log_bm_std` for community biomass to `r r2cond$log_rich_std`
for species richness (Table S1).

```{r}
sem_cap <- "(a) Structural equation models linking the temporal trends of the food-web
structure. The path widths are proportional to the standardized slope
coefficients. The standardized slope coefficients and their associated 95%
confidence intervals are displayed along the path. The marginal R-squared of the
linear models are displayed. (b to f) Direct bivariate relationships derived
from the structural equation model. The distribution of the site temporal trends
are displayed in Fig. S1."
```

@@

```{r}
tar_load(c(semeff_tot, semeff_tot_ok))
semeff_tot <- semeff_tot %>%
  mutate(across(c(response, predictor), ~str_remove(.x, "mean_")))
unique(semeff_tot_ok$predictor)

rdtps <- map(
  setNames(
    c("ct_ff", "w_trph_lvl_avg", "log_bm_std"),
    c("ct", "tlvl", "bm")
    ),
  ~ list(
    bm = get_semeff_ci(resp = .x, term = "log_bm_std"),
    r = get_semeff_ci(resp = .x, term = "log_rich_std")
    )
)
rdsp <- map(
  setNames(
    c("ct_ff", "w_trph_lvl_avg", "log_bm_std"),
    c("ct", "tlvl", "bm")
    ),
  ~ list(
    bm = get_semeff_ci(resp = .x, term = "log_bm_std", model = "spatial"),
    r = get_semeff_ci(resp = .x, term = "log_rich_std", model = "spatial")
    )
)
rdm <- map(
  setNames(
    c("Connectance", "Avg trophic level", "Biomass"),
    c("ct", "tlvl", "bm")
    ),
  ~ list(
    bm = get_semeff_ci(x = semeff_tot_ok, type = "Direct", resp = .x, term = "Biomass", model = "Theoretical model"),
    r = get_semeff_ci(x = semeff_tot_ok, type = "Direct", resp = .x, term = "Species richness", model = "Theoretical model")
    )
)
```

```{r}
ridtps <- map(
  setNames(
    c("ct_ff", "w_trph_lvl_avg", "log_bm_std"),
    c("ct", "tlvl", "bm")
    ),
  ~ list(
    r = get_semeff_ci(resp = .x, term = "log_rich_std", type = "indirect")
    )
)
ridsp <- map(
  setNames(
    c("ct_ff", "w_trph_lvl_avg", "log_bm_std"),
    c("ct", "tlvl", "bm")
    ),
  ~ list(
    r = get_semeff_ci(resp = .x, term = "log_rich_std",
      model = "spatial", type = "indirect")
    )
)
ridm <- map(
  setNames(
    c("Connectance", "Avg trophic level", "Biomass"),
    c("ct", "tlvl", "bm")
    ),
  ~ list(
    bm = get_semeff_ci(x = semeff_tot_ok, type = "Indirect", resp = .x, term = "Biomass", model = "Theoretical model"),
    r = get_semeff_ci(x = semeff_tot_ok, type = "Indirect", resp = .x, term = "Species richness", model = "Theoretical model")
    )
)
rttps <- map(
  setNames(
    c("ct_ff", "w_trph_lvl_avg", "log_bm_std"),
    c("ct", "tlvl", "bm")
    ),
  ~ list(
    bm = get_semeff_ci(resp = .x, term = "log_bm_std", type = "total"),
    r = get_semeff_ci(resp = .x, term = "log_rich_std", type = "total")
    )
)
rtsp <- map(
  setNames(
    c("ct_ff", "w_trph_lvl_avg", "log_bm_std"),
    c("ct", "tlvl", "bm")
    ),
  ~ list(
    bm = get_semeff_ci(resp = .x, term = "log_bm_std",
      model = "spatial", type = "total"),
    r = get_semeff_ci(resp = .x, term = "log_rich_std",
      model = "spatial", type = "total")
    )
)
rtm <- map(
  setNames(
    c("Connectance", "Avg trophic level", "Biomass"),
    c("ct", "tlvl", "bm")
    ),
  ~ list(
    bm = get_semeff_ci(x = semeff_tot_ok, type = "Total", resp = .x, term = "Biomass", model = "Theoretical model"),
    r = get_semeff_ci(x = semeff_tot_ok, type = "Total", resp = .x, term = "Species richness", model = "Theoretical model")
    )
)
```

Using structural equation models to disentangle the direct and indirect effects
of temporal trends in fish species richness and community biomass on temporal
trends in food-web structure, we found a strong positive effect of the temporal
trends in species richness on community biomass ($r_\delta =$ `r rdtps$bm$r`,
Fig. \@ref(fig:sem) a and b, $r_\delta$: standardized slope coefficient, [95%
CI]), meaning that both temporal trends had the same direction in average. In
other word, food-webs that have experienced a decrease in species richness over
time have decreased in community biomass, and conversely, those that increased
in species richness have increased in community biomass.
Temporal trends in community biomass had in turn a positive effect on connectance and
average trophic level (resp. $r_\delta =$ `r rdtps$ct$bm` and `r rdtps$tlvl$bm`,
Fig. \@ref(fig:sem)a, c, e), meaning that food-web whose community biomass
decreased over time have decreased in connectance and average trophic level, and
conversely, those that have experienced an increase in biomass have increased in
connectance and average trophic level. We further found that the temporal
trends of community biomass were positively associated to the temporal trends of
both number and proportion of piscivorous species present (Fig. S3a), meaning
biomass trends are linked to the presence of high trophic level trophic
species.

We further tested if we find the same patterns using spatial variations in
food-web structure rather than temporal changes over time, i.e. if the
space-for-time hypothesis hold. To do so, we used the same dataset and same
structural equation model but with the median food-web structure per site
instead of the temporal trends. We additionally performed theoretical
simulations with a bioenergetic food-web model to further get theoretical
expectations about how food-web structure changes according to community biomass and
species richness variations (see Methods). As in temporal data, species richness
had a positive effect on community biomass both in empirical data and in
theoretical simulations (resp. $r_\delta =$ `r rdsp$bm$r` and `r rdm$bm$r`, Fig.
\@ref(fig:std)). In both empirical data and theoretical simulations, community
biomass had a positive effect on connectance (resp. $r_\delta =$ `r rdsp$ct$bm`
and `r rdm$ct$bm`, Fig. \@ref(fig:std)) and average trophic level (resp.
$r_\delta =$ `r rdsp$tlvl$bm` and `r rdm$tlvl$bm`), as in temporal analysis.

The effects of species richness were more complex, because species richness had
indirect effects on food-web structure through community biomass. Temporal
trends in species richness had a direct negative effect on connectance
($r_\delta =$ `r rdtps$ct$r`, Fig. \@ref(fig:sem)d), but had also a positive
indirect effect through community biomass ($r_\delta =$ `r ridtps$ct$r`), so
that species richness temporal trends had no effect in total on temporal in
connectance ($r_\delta =$ `r rttps$ct$r`, Fig. \@ref(fig:sem)d).
In contrast, temporal trends in species had no direct effects on the temporal
trends in average trophic level ($r_\delta =$ `r rdtps$tlvl$r`, Fig.
\@ref(fig:sem)f), however it had a positive indirect effect through community
biomass ($r_\delta =$ `r ridtps$tlvl$r`). It means food-webs that have
experienced a decrease in species richness over time have decreased in average
trophic level, and conversely, those that increased in species richness have
increased in average trophic level, through community biomass. Furthermore, we
found that temporal trends of the proportion of piscivorous species was
negatively associated with the temporal trends of species richness (Fig. S3),
meaning that communities that increase in richness over time have an increase in
the proportion of non-piscivorous species over time and vice versa.

The effects of spatial variations in species richness on food-web structure
contrasts with the ones found with temporal trends. We found a direct negative
effect of species richness on connectance both using data and model simulations
(resp. $r_\delta =$ `r rdsp$ct$r` and `r rdm$ct$r`, Fig. \@ref(fig:std)) and on
average trophic level (resp. $r_\delta =$ `r rdsp$tlvl$r` and `r rdm$tlvl$r`).
Despite indirect positive effect of species richness through community biomass,
species richness had strong total negative effects on connectance in empirical
data and simulations (resp. $r_\delta =$ `r rtsp$ct$r` and `r rtm$ct$r` ) and
average trophic level (resp. $r_\delta =$ `r rtsp$tlvl$r` and `r rtm$tlvl$r`).

@@

```{r}
std_cap <- paste0(
  "Assessment of the space-for-time hypothesis. Standardized coefficients
  ($r_\\delta$) derived from the spatial and temporal Structural Equation
  Models. The left, middle and right panels display respectively the direct,
  indirect and total effects of community biomass and species richness on the
  Average trophic level (top panel), the connectance (middle panel) and the
  community biomass (bottom panel) The lines display the confidence intervals at
  95%.")
```

@@

We found remarkable heterogeneity in community biomass and species richness
changes over the last decades in riverine fish communities, linked to changes in
food-web structure. Community biomass changes spanned from `r grbmn`% to `r
grbmx`% per decade, species richness from `r grrhn`% to +`r grrhx`% per decade,
in line previous reports of the heterogeneity of local biodiversity changes in
terrestrial, marine and freshwater realms[@dornelas_assemblage_2014;
@blowes_geography_2019; @danet_past_2024]. In turn, connectance temporal changes
spanned from `r grctn` to +`r grctx` per decade and average trophic level, from
`r grtln` to +`r grtlx` per decade. We questioned the effects of community
biomass and species richness changes over time on temporal changes in food-web
structure, that we also tested across space and with simulations using a
bioenergetic food-web model. We found positive effects of species richness
changes on changes in community biomass that were consistent across temporal and
spatial analysis and with the theoretical model. The effects of community
biomass changes on food-web structure were also consistent across analysis,
suggesting that loss of community biomass are linked to the loss of the trophic
species of high trophic level. However, changes in species richness had
contrasting effects on food-web structure depending on the type of analysis,
temporal or spatial and theoretical simulations. It suggests that food-web
structure changes induced by changes in species richness differs from the one
induced by the changes in community biomass, and that they imply different
processes across time and space. In what follows, we discuss how changes
in community biomass and species richness changes over time and space are
related to food-web assembly and disassembly processes.

@@

When considering spatial analysis and model simulations, species richness
effects on food-web structure were highly distinct from the temporal analysis.
Spatial analysis and theoretical models suggest that increase in species
richness results in lower connectance and average trophic level, a well known
patterns in empirical food-web [@dunne_network_2006; @danet_species_2021]. This
result is also in line with other studies which found a mismatch between spatial
and temporal trends [@adler_contrasting_2007; @johnson_testing_2008]. This
discrepancy in space and time is due to the strong assumption that variations in
space reflect differences in assembly stages of ecological communities and then
mirror their temporal assembly/disassembly trajectories. Temporal changes in
species richness might relate to short-term dynamics while spatial gradients
reflect aggregated dynamics over a longer time scale including a far number of
processes including high species turnover along spatial gradients. Our results
then suggest community biomass changes in space might be a good surrogate to
infer assembly and disassembly of food-web in time, but that it is probably not
the case for species richness.

*Final message*

It suggests that the temporal assembly and
disassembly of food-webs are more sensitive to variations in community biomass
than in species richness, in other words that food-web structure is more robust
to changes in species richness than in biomass over time. Our study then
suggests that the changes of food-web structure over time related to changes in
species richness could be actually mediated by changes in community biomass.

@@

We investigated how food-web assemble and disassemble according to changes in
community biomass and species richness. Our study shows that temporal changes in
community biomass induced changes in food-web structure, such as decline in
biomass translates in food-webs with fewer trophic links and top predators while
increases in biomass results in food-webs with more trophic links and top
predators. Species richness temporal trends translate in low changes in food-web
structure, suggesting that species richness changes relates more to trophic role
redundancy than in changes in food-web structure. Using replicating the analysis
over spatial gradient and with a theoretical model, our study shows that
community biomass changes over space and theoretical model results in the same
changes in food-web structure than in temporal analysis, which is not the case
for species richness. It suggests that using space-for-time substitution to
infer assembly and disassembly processes in food-web structure works well for
community biomass and not species richness.

@@

Our study opens stimulating perspectives to further understand of biodiversity
changes alters community structure and ecosystem functioning. It is particularly
needed for ecological networks, whose structure has tremendous effects on
ecosystem functioning and stability. It implies to further investigate which
processes drive the differences in the effects of species richness
changes on food-web structure in space and time.

>>>>

```{r, results=FALSE}
tar_load(glob_tps_trends_decade_no_drivers)
gbm <- get_effect_ci(resp = "log_bm_std", ci_lvl = "level:0.80", p = TRUE)
gr <- get_effect_ci(resp = "log_rich_std", ci_lvl = "level:0.80", p = TRUE)
gt <- get_effect_ci(resp = "w_trph_lvl_avg", ci_lvl = "level:0.80", p = FALSE)
gct <- get_effect_ci(resp = "connectance", ci_lvl = "level:0.80", p = FALSE)
gl <- get_effect_ci(resp = "prop_redundant_links", ci_lvl = "level:0.80", p = FALSE)
```

@@

```{r}
tar_load(summary_pred_station_tps10)
dfv <- function(
  x = summary_pred_station_tps10,
  r = "connectance",
  v = "min",
  d = 4
) {
  out <- x[x$response == r, ][[v]]
  format(round(out, d), nsmall = d)
}
grbmn <- dfv(r = "log_bm_std", v = "min", d = 0)
grbmx <- dfv(r = "log_bm_std", v = "max", d = 0)
grrhn <- dfv(r = "log_rich_std", v = "min", d = 0)
grrhx <- dfv(r = "log_rich_std", v = "max", d = 0)
grctn <- dfv(r = "connectance", v = "min", d = 2)
grctx <- dfv(r = "connectance", v = "max", d = 2)
grtln <- dfv(r = "w_trph_lvl_avg", v = "min", d = 2)
grtlx <- dfv(r = "w_trph_lvl_avg", v = "max", d = 2)
grln <- dfv(r = "prop_redundant_links", v = "min", d = 2)
grlx <- dfv(r = "prop_redundant_links", v = "max", d = 2)
```

Using a Bayesian hierarchical model to jointly estimate the temporal trends of
food-web structure at all sites, we found weak evidence of an overall decrease
in community biomass (CI80%:`r gbm` by decade; median [Credible Interval], Fig.
S1) but no evidence of an overall temporal increase in species richness
(CI80%: `r gr` by decade) across the `r nb_st` sites. We did not find evidence
of average temporal trends in connectance and average trophic level (CI80%:
`r gct` and `r gt` respectively), but an increase in redundancy (measured by the
proportion of redundant links, CI80%: `r gl`). However, we found a great
heterogeneity in the temporal trends estimated at the site level (community
biomass: from `r grbmn`% to `r grbmx`% per decade, species richness: `r grrhn`
to +`r grrhx`%, average trophic level: `r grtln` to +`r grtlx`, connectance: `r
grctn` to +`r grctx`, redundancy: `r grln` to +`r grlx`, Fig. S1) which was well
captured by the hierarchical model with conditional R2 going from `r r2cond$log_bm_std` for community biomass to `r r2cond$log_rich_std` for species
richness (Table S1).

```{r}
sem_cap <- "(a) Structural equation models linking the temporal trends of the food-web
structure. The path widths are proportional to the standardized slope
coefficients. The standardized slope coefficients and their associated 95%
confidence intervals obtained by bootstrapping are displayed along the path. The
marginal R-squared of the linear models are displayed. (b to f) Direct bivariate
relationships derived from the structural equation model. The distribution of
the site temporal trends are displayed in Fig. S1. All the effects are reported
in Table S4."
```

@@

```{r}
tar_load(c(semeff_tot, semeff_tot_ok))
semeff_tot <- semeff_tot %>%
  mutate(across(c(response, predictor), ~str_remove(.x, "mean_")))

rdtps <- map(
  setNames(
    c("connectance", "w_trph_lvl_avg", "log_bm_std", "prop_redundant_links"),
    c("ct", "tlvl", "bm", "red")
    ),
  ~ list(
    bm = get_semeff_ci(resp = .x, term = "log_bm_std"),
    r = get_semeff_ci(resp = .x, term = "log_rich_std")
    )
)
rdsp <- map(
  setNames(
    c("connectance", "w_trph_lvl_avg", "log_bm_std", "prop_redundant_links"),
    c("ct", "tlvl", "bm", "red")
    ),
  ~ list(
    bm = get_semeff_ci(resp = .x, term = "log_bm_std", model = "spatial"),
    r = get_semeff_ci(resp = .x, term = "log_rich_std", model = "spatial")
    )
)
rdm <- map(
  setNames(
    c("Connectance", "Avg trophic level", "Biomass", "Redundancy"),
    c("ct", "tlvl", "bm", "red")
    ),
  ~ list(
    bm = get_semeff_ci(x = semeff_tot_ok, type = "Direct", resp = .x, term = "Biomass", model = "Theoretical model"),
    r = get_semeff_ci(x = semeff_tot_ok, type = "Direct", resp = .x, term = "Species richness", model = "Theoretical model")
    )
)
```

```{r}
ridtps <- map(
  setNames(
    c("connectance", "w_trph_lvl_avg", "log_bm_std", "prop_redundant_links"),
    c("ct", "tlvl", "bm", "red")
    ),
  ~ list(
    r = get_semeff_ci(resp = .x, term = "log_rich_std", type = "indirect")
    )
)
ridsp <- map(
  setNames(
    c("connectance", "w_trph_lvl_avg", "log_bm_std", "prop_redundant_links"),
    c("ct", "tlvl", "bm", "red")
    ),
  ~ list(
    r = get_semeff_ci(resp = .x, term = "log_rich_std",
      model = "spatial", type = "indirect")
    )
)
ridm <- map(
  setNames(
    c("Connectance", "Avg trophic level", "Biomass", "Redundancy"),
    c("ct", "tlvl", "bm", "red")
    ),
  ~ list(
    bm = get_semeff_ci(x = semeff_tot_ok, type = "Indirect", resp = .x, term = "Biomass", model = "Theoretical model"),
    r = get_semeff_ci(x = semeff_tot_ok, type = "Indirect", resp = .x, term = "Species richness", model = "Theoretical model")
    )
)
rttps <- map(
  setNames(
    c("connectance", "w_trph_lvl_avg", "log_bm_std", "prop_redundant_links"),
    c("ct", "tlvl", "bm", "red")
    ),
  ~ list(
    bm = get_semeff_ci(resp = .x, term = "log_bm_std", type = "total"),
    r = get_semeff_ci(resp = .x, term = "log_rich_std", type = "total")
    )
)
rtsp <- map(
  setNames(
    c("connectance", "w_trph_lvl_avg", "log_bm_std", "prop_redundant_links"),
    c("ct", "tlvl", "bm", "red")
    ),
  ~ list(
    bm = get_semeff_ci(resp = .x, term = "log_bm_std",
      model = "spatial", type = "total"),
    r = get_semeff_ci(resp = .x, term = "log_rich_std",
      model = "spatial", type = "total")
    )
)
rtm <- map(
  setNames(
    c("Connectance", "Avg trophic level", "Biomass", "Redundancy"),
    c("ct", "tlvl", "bm", "red")
    ),
  ~ list(
    bm = get_semeff_ci(x = semeff_tot_ok, type = "Total", resp = .x, term = "Biomass", model = "Theoretical model"),
    r = get_semeff_ci(x = semeff_tot_ok, type = "Total", resp = .x, term = "Species richness", model = "Theoretical model")
    )
)
```

Using structural equation models to disentangle the direct and indirect effects
of temporal trends in fish species richness and community biomass on temporal
trends in food-web structure, we found a strong positive effect of the temporal
trends in species richness on community biomass ($r_\delta =$ `r rdtps$bm$r`,
Fig. \@ref(fig:sem) a and b, $r_\delta$: standardized slope coefficient, [95%
CI]), meaning that both temporal trends had the same direction in average. In
other word, food-webs that have experienced a decrease in species richness over
time have decreased in community biomass, and conversely, those that increased
in species richness have increased in community biomass.
Temporal trends in community biomass had in turn a positive effect on connectance,
average trophic level, and redundancy (resp. $r_\delta =$ `r rdtps$ct$bm`,
`r rdtps$tlvl$bm`, `r rdtps$red$bm`, Fig. \@ref(fig:sem)b, c, d), meaning that
food-web whose community biomass decreased over time have decreased in
connectance, average trophic level and redundancy, and
conversely, those that have experienced an increase in biomass have increased in
connectance, average trophic level and redundancy. We further found that the temporal
trends of community biomass were positively associated to the temporal trends of
both number and proportion of piscivorous species present (Fig. S3a), meaning
biomass trends are linked to the presence of high trophic level trophic
species.

We further tested if we find the same patterns using spatial variations in
food-web structure rather than temporal changes over time, i.e. if the
space-for-time hypothesis hold. To do so, we used the same dataset and same
structural equation model but with the median food-web structure per site
instead of the temporal trends. We additionally performed theoretical
simulations with a bioenergetic food-web model to further get theoretical
expectations about how food-web structure changes according to community biomass and
species richness variations (see Methods). As in temporal data, species richness
had a positive effect on community biomass both in empirical data and in
theoretical simulations (resp. $r_\delta =$ `r rdsp$bm$r` and `r rdm$bm$r`, Fig.
\@ref(fig:std) and S5). In both empirical data and theoretical simulations, community
biomass had a positive effect on connectance (resp. $r_\delta =$ `r rdsp$ct$bm`
and `r rdm$ct$bm`, Fig. \@ref(fig:std)), average trophic level (resp.
$r_\delta =$ `r rdsp$tlvl$bm` and `r rdm$tlvl$bm`), and redundancy (resp.
$r_\delta =$ `r rdsp$red$bm` and `r rdm$red$bm`), as in temporal analysis.

The effects of species richness were more complex, because species richness had
indirect effects on food-web structure through community biomass. Temporal
trends in species richness had a direct negative effect on connectance
($r_\delta =$ `r rdtps$ct$r`, Fig. \@ref(fig:sem)f), a positive
indirect effect through community biomass ($r_\delta =$ `r ridtps$ct$r`), so
that species richness temporal trends had a negative effect in total on temporal in
connectance ($r_\delta =$ `r rttps$ct$r`, Fig. \@ref(fig:std)).
In contrast, temporal trends in species had no direct effects on the temporal
trends in average trophic level ($r_\delta =$ `r rdtps$tlvl$r`, Fig.
\@ref(fig:sem)g), a positive indirect effect through community
biomass ($r_\delta =$ `r ridtps$tlvl$r`), which adds up to a positive total
effect of species richness temporal trends on average trophic level ($r_\delta
=$ `r rttps$tlvl$r`, Fig. \@ref(fig:std)). Species richness temporal trends had
direct and indirect positive effect on the temporal trends in redundancy (resp.
$r_\delta =$ `r rdtps$red$r`, `r ridtps$red$r`), which adds up to a strong
positive effect $r_\delta =$ `r rttps$red$r`. Furthermore, we found that
temporal trends of the proportion of piscivorous species was negatively
associated with the temporal trends of species richness (Fig. S3), meaning that
communities that increase in richness over time have an increase in the
proportion of non-piscivorous species over time and vice versa. It means
food-webs that have experienced a decrease in species richness over time have
decreased in average trophic level and redundancy but increase in connectance,
and conversely, those that increased in species richness have increased in
average trophic level, redundancy, but decreased in connectance.

The effects of spatial variations in species richness on food-web structure
mostly agreed with the ones found with temporal trends. We found a direct negative
effect of species richness on connectance both using data and model simulations
(resp. $r_\delta =$ `r rdsp$ct$r` and `r rdm$ct$r`, Fig. \@ref(fig:std)), on
average trophic level (resp. $r_\delta =$ `r rdsp$tlvl$r` and `r rdm$tlvl$r`),
but a direct positive effect on redundancy (resp. $r_\delta =$ `r rdsp$red$r`
and `r rdm$red$r`). Species richess had indirect positive effect of species
richness through community biomass, which added up to strong total negative
effects on connectance in empirical data and simulations (resp. $r_\delta =$
`r rtsp$ct$r` and `r rtm$ct$r` ) and positive on redundancy (resp. $r_\delta =$
`r rtsp$red$r` and `r rtm$red$r`). We found that species richness had a total
positive effect on average trophic level in bioenergetic model, in agreement
with temporal trends (resp. $r_\delta =$ `r rtm$tlvl$r`), but a total negative
effect in spatial variations (resp. $r_\delta =$ `r rtsp$tlvl$r`).

@@

```{r}
std_cap <- paste0(
  "Direct and total effects of community biomass and species richness.
  Standardized coefficients ($r_\\delta$) derived from the temporal, spatial,
  and bioenergetic model Structural Equation Models. Points and lines display
  the estimated effects and their confidence intervals at 95\\%. Values are
  displayed in Table SXX, R-squared in Table SXX, and the visualisation of the
  bivariate relationships for the spatial and bioenergetic model SEMs are
  displayed in Figure SXX.")

```

@@

We found remarkable heterogeneity in community biomass and species richness
changes over the last decades in riverine fish communities, linked to changes in
food-web structure. Community biomass changes spanned from `r grbmn`% to `r
grbmx`% per decade, species richness from `r grrhn`% to +`r grrhx`% per decade,
in line previous reports of the heterogeneity of local biodiversity changes in
terrestrial, marine and freshwater realms[@dornelas_assemblage_2014;
@blowes_geography_2019; @danet_past_2024]. In turn, connectance temporal changes
spanned from `r grctn` to +`r grctx` per decade and average trophic level, from
`r grtln` to +`r grtlx` per decade. We questioned the effects of community
biomass and species richness changes over time on temporal changes in food-web
structure, that we also tested across space and with simulations using a
bioenergetic food-web model. We found positive effects of species richness
changes on changes in community biomass that were consistent across temporal and
spatial analysis and with the theoretical model. The effects of community
biomass changes on food-web structure were also consistent across analysis,
suggesting that loss of community biomass are linked to the loss of the trophic
species of high trophic level. However, changes in species richness had
contrasting effects on food-web structure depending on the type of analysis,
temporal or spatial and theoretical simulations. It suggests that food-web
structural changes induced by changes in species richness differs from the one
induced by the changes in community biomass, and that they imply different
processes across time and space. However, all analysis converged to indicate
that community biomass and species richness declines trigger a decrease in
food-web redundancy. In what follows, we discuss how changes in community
biomass and species richness changes over time and space are related to food-web
assembly and disassembly processes.

@@

When considering spatial analysis and model simulations, species richness
effects on food-web structure mostly consistent from the temporal analysis. All
analysis converged to indicate that increases in species richness result in
decreases in connectance, a well documented pattern in empirical food-webs
[@dunne_network_2006; @danet_species_2021]. However, spatial analysis found that
species richness decrease average trophic level, while temporal analysis and
theoretical models found that species richness increases result in higher
average trophic level. This result is also in line with other studies which
found a mismatch between spatial and temporal trends [@adler_contrasting_2007;
@johnson_testing_2008]. This discrepancy in space and time is due to the strong
assumption that variations in space reflect differences in assembly stages of
ecological communities and then mirror their temporal assembly/disassembly
trajectories. Temporal changes in species richness might relate to short-term
dynamics while spatial gradients reflect aggregated dynamics over a longer time
scale including a far number of processes including high species turnover along
spatial gradients. Our results then suggest community biomass changes in space
might be a good surrogate to infer assembly and disassembly of food-web in time,
but that it is probably not always the case for species richness.

@@

- Revise: biomass pyramids, connectance different from top predators
- Add redundancy

We investigated how food-web assemble and disassemble according to changes in
community biomass and species richness. Our study shows that temporal changes in
community biomass induced changes in food-web structure, such as decline in
biomass translates in food-webs with *lower connectance* and fewer top predators
while increases in biomass results in food-webs with higher connectance and more
top predators. In turn, decline in species richness translates in food-webs with
fewer trophic links and top predators while
increases in biomass results in food-webs with more trophic links and top
predatorstranslate in low changes in
food-web structure, suggesting that species richness changes relates more to
trophic role
redundancy than in changes in food-web structure. Using replicating the analysis
over spatial gradient and with a theoretical model, our study shows that
community biomass changes over space and theoretical model results in the same
changes in food-web structure than in temporal analysis, which is not the case
for species richness. It suggests that using space-for-time substitution to
infer assembly and disassembly processes in food-web structure works well for
community biomass but not so well for species richness.

@@
